---
title: "Distributing jobs for high-performance computing (HPC) clusters (e.g., via Slurm)"
author: "Phil Chalmers"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: false
    number_sections: true 
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: false
      smooth_scroll: false
vignette: >
  %\VignetteIndexEntry{HPC cluster array jobs (e.g., via Slurm)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r nomessages, echo = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.height = 5,
  fig.width = 5
)
options(digits=4)
par(mar=c(3,3,1,1)+.1)
```

```{r include=FALSE}
library(SimDesign)
```


# Introduction

The purpose of this vignette is to demonstrate how to utilize `SimDesign` in the context of distributing many jobs across independent computing environments, such as high-performance computing (HPC) clusters, in a way that allows for reproducibility of the simulation conditions, resubmission of jobs in case of incomplete collection of results within a specified HPC execution time budget, and to ensure that random number generation across the entire simulation (and subsequent resubmissions, if required) are properly manged throughout given the batch nature of the job. The following text and examples are primarily for managing larger simulations, often with thousands of replications over many simulation conditions (i.e., rows in the defined `design` object)  which generally require non-trivial amounts of computing resources to execute (hence, the need for super-computing resources and job schedulers like Slurm, TORQUE, MAUI, among others). 

For information about Slurm's Job Array support in particular, which this vignette uses as an example, see https://slurm.schedmd.com/job_array.html 

# Standard setup on HPC cluster

To start, the structure of the simulation code used later on to distribute the jobs to the HPC scheduler is effectively the same as the usual generate-analyse-summarise workflow described in `runSimulation()`, with a few organizational exceptions. As such, this is always a good place to start when designing, testing, and debugging a simulation experiment before submitting to HPC clusters.

**IMPORTANT: Only after the vast majority of the bugs and coding logic have been work out should you consider moving on to the next step involving HPC clusters**. If your code is not well vetted in this step then any later jobs evaluated on the HPC cluster will be a waste of time and resources (garbage-in, garbage-out).

### Example

Suppose the following simulation was to be evaluated, though for time constraint reasons would not be possible to execute on a single computer (or a smaller network of computers) and therefore should be submitted to an HPC cluster. 

The following script (hypothetically written in a file called `SimDesign_simulation.R`) contains a simulation experiment whose instructions are to be submitted to the Slurm scheduler. To do so, the `sbatch` utility is used along with the set of instructions specifying the type of hardware required in the file `slurmInstructions.slurm`. In the R side of the simulation, the defined code must grab all available cores (minus 1) that are detectable via `parallel::detectCores()`, which occurs automatically when using `runSimulation(..., parallel=TRUE)`. 
```{r}
# SimDesign::SimFunctions()
library(SimDesign)

Design <- createDesign(N = c(10, 20, 30))

Generate <- function(condition, fixed_objects = NULL) {
    dat <- with(condition, rnorm(N, 10, 5)) # distributed N(10, 5)
    dat
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(mean=mean(dat), median=median(dat)) # mean/median of sample data
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL){
    colMeans(results)
}
```

```{r eval=FALSE}
# standard setup (not ideal for HPC clusters as parallelization
#  occurs within the design conditions, not across)
res <- runSimulation(design=Design, replications=10000, generate=Generate,
                     analyse=Analyse, summarise=Summarise, parallel=TRUE, 
                     filename='mysim')
```

In the standard `runSimulation(..., parallel=TRUE)` setup the 10,000 
replications would be distributed to the available computing cores and evaluated
independently across the three row conditions in the `design` object. However, this process is only
executed in sequence: `design[1, ]` is evaluated first and, only after the 10,000 replications 
are collected, `design[2, ]` is evaluated until it is complete, then `design[3, ]`, and so on. 

As well, in order for this approach to be at all optimal the HPC cluster must assign a job containing a very large number of resources; specifically, higher RAM and CPUs. To demonstrate, in the following `slurmInstructions.slurm` file a larger number of CPUs are requested when allocating the computational structure/cluster associated with this job, as well as larger amounts of RAM. 

```
#!/bin/bash
#SBATCH --job-name="My simulation (multiple CPUs)"
#SBATCH --mail-type=ALL
#SBATCH --mail-user=somewhere@out.there
#SBATCH --output=/dev/null    ## (optional) delete .out files
#SBATCH --time=12:00:00       ## HH:MM:SS
#SBATCH --mem=192G            ## Build a computer with 192GB of RAM 
#SBATCH --cpus-per-task=96    ## Build a computer with 96 cores

module load R/4.3.1
Rscript --vanilla SimDesign_simulation.R
```

This job request a computing cluster be built with 192 GB of RAM with 96 CPUs, which the `SimDesign_simulation.R` is evaluated in, and is submitted to the scheduler via `sbatch slurmInstructions.slurm`.

### Limitations

While generally effective at distributing the computational load, there are a few limitations of the above approach:

1) For simulations with varying execution times this will create a great deal of resource waste.
    - Due to the row-evaluation nature of the `design` conditions, computing cores will at some point sit idle while waiting for the remaining experiments to complete their job. This occurs for each row in the `design` input (i.e., per simulation condition)
    - As such, simulation experiments with many conditions to evaluate will suffer most due to the rolling overhead, resulting in wasted resource management (not kind to other users of the HPC cluster) and ultimately results in jobs that take longer to complete
2) Designers must first estimate the number of *total* amount of CPUs/RAM/time required, while being as conservative as possible
    - Problematic because fewer resource jobs are given higher priority in the scheduler, and because you may be taking resources away from other researchers if the cores sit idle
3) Parallel distribution across the allocated resources (e.g., across two nodes, both with 48 cores) incurs some overhead that grows as a function of the size of the defined cluster 
    - Using a 96 CPU cluster will not result in a 96x speedup of a 1 CPU job. In fact, the larger the allocated cluster, the worse the performance efficiency
4) The scheduler must wait until all resources (e.g., RAM and CPUs) *simultaneously* become available to allocate the requested specifications, which can take a good amount of time to allocate if the resources requested are excessively high (hence, become low *priority* on the scheduler)
    - If you request 96 CPUs with 192 GB of RAM then this will take considerably longer to allocate compared to requesting 96 independent computing arrays with 1 CPU and 2 GB of RAM (the latter approach is described in the next section)
    - In concert with point 3), this results in jobs that a) take longer to get started as they will sit longer in the queue and b) may not distribute the load efficiently enough, thereby resulting in larger wall time then should have been necessary 
5) Submitting independent multiple batches to the cluster makes it more difficult to guarantee the quality of the random numbers
    - Setting the `seed` for each condition ensure that within each `design` condition the random numbers are high quality, however there is no guarantee that repeated use of `set.seed()` will result in high-quality random numbers (see next section for example)
    - Hence, repeated job submissions of this type, even with unique seeds per condition, may not generate high quality numbers if repeated too many times (alternative is to isolate each `design` row and submit each row as a unique job, which is demonstrated near the end of this vignette)
6) Finally, and perhaps most problematic in simulation experiment applications, schedulers frequently cap the maximum number of resources that can be requested (e.g., 256 GB of RAM, 200 CPUs), which limits the application of large RAM and CPU jobs
    - Note that to avoid wasting time by swapping/paging, schedulers will never allocate jobs whose memory requirements exceed the amount of available memory 
    
To address these and other computational inefficiencies/wasted resources, one can instead switch from the cluster-based approach above to an *array* distribution approach, discussed in the next section. 

***

# Array jobs 

For HPC computing it is often more optimal to distribute both replications *and* conditions simultaneously to
unique computing nodes (termed **arrays**) to effectively break the problem in several mini-batches (e.g., split the simulation into 1000+ independent pieces, and collect the results later). 
As such, the above `design` object and `runSimulation()` structure above does not readily lend itself to optimal distribution for the array scheduler to manage. Nevertheless, the 
core components are still useful for initial code design, testing, and debugging, and therefore serve as a necessary first step when writing simulation experiment code prior to submitting to an HPC cluster.

After defining and testing your simulation to ensure that it works as expected,
it now comes the time to setup the components required for organizing the HPC
cluster submission using the `runArraySimulation()` function. 

## Converting `runSimulation()` workflow to one for `runArraySimulation()`

The job of `runArraySimulation()` is to utilize the relevant information defined in the `.sh` or `.slurm` script. This is done by extracting information provided by the scheduler (specifically, via an `arrayID`), which is used to select specific subsets of the `design` rows. However, unlike `runSimulation()` the function `runArraySimulation()` has been designed to control important information pertaining to `.Random.seeds` and other relevant distribution information that allow for the rows in the `design` object itself to contain **repeated experimental condition information**. This allows both the `design` rows *and* `replication` information to be optimally distributed to the HPC cluster.

The following example presents the essential modifications required to move from a single `runSimulation()` workflow to a batch workflow suitable for `runArraySimulation()` and Slurm scheduler.

## Expand the standard simulation `design` object for each array ID 

Suppose that 300 computing cores were independently available on the HPC cluster, though 
the availability of these cores only trickle in as a function of the schedulers 
decided availability. In this case, the strategy would be to split up the `3 * 10000 = 30000` condition-by-replications experiments across the gradually available resources, where jobs are evaluated in parallel and at different times.

Given the above specifications, you may decide that each of the 300 computing nodes requested to the scheduler should evaluate exactly 100 replications each (`nrow(design) * 10000 / 300 = 100`). In this case, `expandDesign()` is used to repeat each row condition 100 times, creating a new *expanded* `design` object with 300 rows instead of 3. To accommodate for the new rows, the associated `replications` should now be defined according to how many replications should be used within each of these to-be-distributed conditions, which need not be equal (see below).

```{r}
rc <- 100   # number of times the design row was repeated
Design300 <- expandDesign(Design, repeat_conditions = rc)
Design300

# target replication number for each condition
rep_target <- 10000

# replications per row in Design300
replications <- rep(rep_target  / rc, nrow(Design300))
```
The above approach assumes that each `design` condition is equally balanced in terms of computing time and resources, though if this is not the case (e.g., the last condition contains notably higher computing times than the first two conditions) then `repeat_conditions` can be specified as a vector instead, such as `repeat_conditions = c(100, 100, 1000)`, which for the latter portion would be associated with a 10 replications per distributed node instead of 100. 

```{r}
rc <- c(100, 100, 1000)
DesignUnbalanced <- expandDesign(Design, repeat_conditions = rc)
DesignUnbalanced

rep_target <- 10000
replicationsUnbalanced <- rep(rep_target / rc, times = rc)
head(replicationsUnbalanced)
table(replicationsUnbalanced)
```

Regardless of whether the expanded design is balanced or unbalanced each row in the resulting expanded design object will be assigned to a unique computing array node, identified according to the Slurm assigned array ID. 

## Construct and record proper random seeds

In principle, the expanded `Design300` object above could be passed as `runSimulation(Design300, replications=100, ...)` to evaluate each of the repeated conditions, where each row is now replicated only 100 times. However, there is now an issue with respect to the random seed management in that use of functions such as `set.seed()` and friends are no longer viable. This is because repeated use of `set.seed()` does not itself guarantee independent high-quality random numbers between different instances. For example,

```{r}
set.seed(0)
x <- runif(100)
set.seed(1)
y <- runif(100)

plot(x, y)           ## seemingly independent
plot(x[-1], y[-100]) ## subsets perfectly correlated
```

This issue is generally not problem in the standard `runSimulation()` approach as within each design condition high quality random numbers are used by default, and any potentially repeated number sequences *across* the conditions are highly unlikely to affect the quality of the overall simulation experiment (the conditions themselves typically generate and manage random numbers in different ways due to the varying simulation factors, such as sample sizes, variance conditions, fitted models, number of variables, type of probability distributions use, and so on). However, in the `expandDesign()` setup the likelihood of witnessing correlated/redundant random samples increases very quickly, which is particularly problematic within each distributed replication set; hence, special care must be taken to ensure that proper seeds are distributed to each job array.

Fortunately, seeds are easy to manage with the `genSeeds()` function using a two-step approach, which is internally managed by `runArraySimulation()` by supplying an initial seed (`iseed`) value and the associated array ID (`arrayID`). Doing so will utilize L'Ecuyer's (1999) method, which constructs sequentially computed `.Random.seed` states to ensure independence across all replications and conditions. Note that `iseed` must constant across all job arrays, so **make sure to define `iseed` once and only once**!

```{r}
# genSeeds()   # do this once on the main node/home computer and store the number!
iseed <- 1276149341
```

As discussed in the FAQ section at the bottom, this associated value will also allow for generation of new `.Random.seed` elements if (or when) a second or third set of simulation jobs should be submitted to the HPC cluster at a later time but must also generate simulated data that is independent from the initial submission(s).  

## Extract array ID information from the `.slurm` script

When submitting to the HPC cluster you'll need to include information about how the scheduler should distribute the simulation experiment to the workers. In Slurm systems, you may have a script such as the following, stored into a suitable `.slurm` file:

```
#!/bin/bash
#SBATCH --job-name="My simulation (array jobs, distributing conditions + replications)"
#SBATCH --mail-type=ALL
#SBATCH --mail-user=somewhere@out.there
#SBATCH --output=/dev/null    ## (optional) delete .out files
#SBATCH --time=12:00:00       ## HH:MM:SS
#SBATCH --mem=2G
#SBATCH --cpus-per-task=1
#SBATCH --array=1-300         ## Slurm schedulers often allow up to 10,000 arrays

module load R/4.3.1
Rscript --vanilla mySimDesignScript.R
```

For reference later, label this file `simulation.slurm` as this is the file that must be submitted to the scheduler when it's time.

The top part of this `.slurm` file provides the BASH instructions for the Slurm scheduler via the `#SBATCH` statements. In this case, how many array jobs to queue (1 through 300), how much memory to use per job (2GB), time limits (12 hours), and more; [see here for SBATCH details](https://slurm.schedmd.com/sbatch.html).

The most important input to focus on in this context is **#SBATCH --array=1-300** as this is what is used by the Slurm scheduler to assign a unique ID to each array job. What the scheduler does is take the defined `mySimDesignScript.R` script and send this to 300 independent resources (each with 1 CPU and 2GB of RAM, in this case), where the independent jobs are assigned a unique array ID number within the `--array=1-300` range (e.g., distribution to the first computing resource would be assigned `arrayID=1`, the second resource `arrayID=2`, and so on). In the `runArraySimulation()` function this is used to subset the `Design300` object by row; hence, *the array range must correspond to the row identifiers in the `design` object for proper subsetting!* 

Collecting this single number assigned by the Slurm scheduler is also easy. Just include 
```{r eval=FALSE}
# get assigned array ID (default uses type = 'slurm')
arrayID <- getArrayID()
```
to obtain the associated array ID, which is this example will be a single `integer` value between 1 and 300. This value is used in the final execution step via  `runArraySimulation(..., arrayID=arrayID)`, which we finally turn to.

## Organize information for `runArraySimulation()`

With all the prerequisite steps in place we're finally ready to pass all information to `runArraySimulation()`, which is effectively a wrapper to `runSimulation()` that suppresses verbose outputs, takes subsets of the `Design300` object given the supplied `arrayID` (and other objects, such as `replications`, `seeds`, etc), forces evaluation on a single CPU (hence, `#SBATCH --cpus-per-task=1` should be used by default, unless there is further parallelization to occur within the replications, such as via `OpenMP`), manages the random number generation seeds in a tractable way, and saves the `SimDesign` results to file names based on the `filename` argument with suffixes associated with the `arrayID` (e.g., `filename='mysim'` will save the files `mysim-1.rds` for array 1, `mysim-2.rds` for array 2, ..., `mysim-300.rds` for array 300). 

```{r eval=FALSE}
# run the simulation on subset based on arrayID subset information
runArraySimulation(design=Design300, replications=replications,
                   generate=Generate, analyse=Analyse,
                   summarise=Summarise, iseed=iseed,
                   arrayID=arrayID, filename='mysim')
```

And that's it! 

The above will store all the `mysim-#.rds` files in the directory where the job was submitted, which is somewhat on the messy side, so you may also want to specify a directory name to store the simulation files to. Hence,  on the main (i.e., landing) location associated with your `ssh` account create a directory, using something like `mkdir mysimfiles` (or in R, `dir.create('mysimfiles')`) in the location where your `.R` and `.slurm` files are stored. Then the following can be used to store all 300 collected `.rds` files, making use of the `dirname` argument. 

```{r eval=FALSE}
# run the simulation on subset based on arrayID subset information
runArraySimulation(design=Design300, replications=replications,
                   generate=Generate, analyse=Analyse,
                   summarise=Summarise, iseed=iseed, arrayID=arrayID, 
                   dirname='mysimfiles', filename='mysim')
```
Regardless, the hard part is done here, though other information could be included by way of the `control` list input if necessary, such as including explicit time/resource limits in the R executions within array jobs themselves (see the FAQ section for further information).

## Putting it all together

Below is the complete submission script collecting everything that was presented above. This assumes that 

- The `.R` file with the simulation code is stored in the file `mySimDesignScript.R`,
- A suitable Slurm instruction file has been created in the file `simulation.slurm`, which points to `mySimDesignScript.R` and includes the relevant `R` modules, and
- A directory called `mysimfiles/` has been created for storing the files on the computer used to submit the array job


```{r eval=FALSE}
library(SimDesign)

Design <- createDesign(N = c(10, 20, 30))

Generate <- function(condition, fixed_objects = NULL) {
    dat <- with(condition, rnorm(N, 10, 5)) # distributed N(10, 5)
    dat
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(mean=mean(dat), median=median(dat)) # mean/median of sample data
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL){
    colMeans(results)
}

# expand the design to create 300 rows with associated replications
rc <- 100
Design300 <- expandDesign(Design, repeat_conditions = rc)

rep_target <- 10000
replications <- rep(rep_target / rc, nrow(Design300))

# genSeeds() # do this once on the main node/home computer, and store the number!
iseed <- 1276149341

# get assigned array ID (default uses type = 'slurm')
arrayID <- getArrayID()

# run the simulation on subset based on arrayID subset information
runArraySimulation(design=Design300, replications=replications,
                   generate=Generate, analyse=Analyse,
                   summarise=Summarise, iseed=iseed, arrayID=arrayID, 
                   dirname='mysimfiles', filename='mysim')
```

This file is then submitted to the job scheduler via `sbatch`, pointing to the `.slurm` instructions.
```
sbatch simulation.slurm
```
Once complete you can now go get a beer, coffee, or whatever else tickles your fancy to celebrate as the hard part is over.

# Post-evaluation: Combine the files 

After some time has elapsed, and the job evaluation is now complete, you'll have access to the complete set of simulation files store in the file names `mysim-#.rds`. The final step in this process is then to collect all independent results into a simulation object that resembles what would have been returned from the canonical `runSimulation()` function. Fortunately, this is easy to do with `SimCollect()`. All you must do at this point is change to the working directory containing the simulation files and use `SimCollect()` with the files names to be aggregated:


```{r eval=FALSE}
setwd('mysimfiles')
library(SimDesign)
Final <- SimCollect(files=dir())
Final
```

```
# A tibble: 3 × 8
      N    mean  median REPLICATIONS   SIM_TIME  COMPLETED               
<dbl>   <dbl>   <dbl>           <dbl>  <chr>     <chr>                   
1    10  9.9973  9.9934        10000   23.42s    Thu Apr  4 11:50:11 2024
2    20 10.007  10.015         10000   24.24s    Thu Apr  4 11:50:35 2024
3    30 10.003  10.007         10000   24.39s    Thu Apr  4 11:51:00 2024
```

This function detects which `Design300` rows belong to the original `Design` object, collapse the meta-statistic results, and stored `results` information accordingly. No fuss, no mess. Of course, you'll want to store this object for later use as this is the complete collection of the results from the 300 array jobs, organized into one neat little (object) package.

```{r eval=FALSE}
# save the aggregated simulation object for subsequent analyses
saveRDS(Final, "../final_sim.rds")
```
You should now consider moving this `"final_sim.rds"` off the Slurm landing node and onto your home computer via `scp` or your other favourite method (e.g., using `WinSCP` on Windows). You could also move all the saved `*.rds` files off your scp landing in there is need to inspect these files further (e.g., for debugging purposes).

***

# Array jobs and multicore computing simultaneously

Of course, nothing really stops you from mixing and matching the above ideas related to multicore computing and array jobs on Slurm and other HPC clusters. For example, if you wanted to take the original `design` object and submit batches of these instead (e.g., submit one or more rows of the `design` object as an array job), where within each batch multicore processing is requested, then something like the following would work just fine: 

```
#!/bin/bash
#SBATCH --job-name="My simulation (arrays + multiple CPUs)"
#SBATCH --mail-type=ALL
#SBATCH --mail-user=somewhere@out.there
#SBATCH --output=/dev/null    ## (optional) delete .out files
#SBATCH --time=04:00:00       ## HH:MM:SS
#SBATCH --mem=32G             ## 32 GB RAM per array
#SBATCH --cpus-per-task=16    ## 16 CPUs per array, likely built from 1 node
#SBATCH --array=1-9           ## 9 array jobs 

module load R/4.3.1
Rscript --vanilla mySimDesignScript.R
```

with the associated `.R` file containing, in this case, nine simulation conditions across the rows in `Design`.

```{r}
library(SimDesign)

Design <- createDesign(N = c(10, 20, 30),
                       SD = c(1,2,3))

Generate <- function(condition, fixed_objects = NULL) {
    dat <- with(condition, rnorm(N, 10, sd=SD)) # distributed N(10, 5)
    dat
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(mean=mean(dat), median=median(dat)) # mean/median of sample data
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL){
    colMeans(results)
}

Design
```

Depending on the intensity of the conditions, you may choose to distribute more than one row of  the `Design` object to each array (`multirow=TRUE` in the following), otherwise the more natural choice is to distribute each row in the `Design` object to each assigned array.

```{r eval=FALSE}

# get array ID
arrayID <- getArrayID()

multirow <- FALSE  # submit multiple rows of Design object to array?
if(multirow){
    # If selecting multiple design rows per array, such as the first 3 rows, 
    #  then next 3 rows, and so on, something like the following would work
    s <- c(seq(from=1, to=nrow(Design), by=3), nrow(Design)+1L)
    
    ## For arrayID=1, rows2pick is c(1,2,3); for arrayID2, rows2pick is c(4,5,6)
    rows2pick <- s[arrayID]:(s[arrayID + 1] - 1)
    filename <- paste0('mysim-', paste0(rows2pick, collapse=''))
} else {
    # otherwise, submit each row independently across array
    rows2pick <- arrayID
    filename <- paste0('mysim-', rows2pick)
}

# Make sure parallel=TRUE flag is on! Also, it's important to change the computer
# name to something unique to the array job to avoid overwriting files (even temporary ones)
runSimulation(design=Design[rows2pick, ], replications=10000, 
              generate=Generate, analyse=Analyse, summarise=Summarise, 
              parallel=TRUE, filename=filename, 
              save_details=list(compname=paste0('array-', arrayID)))  
```

When complete, the function `SimCollect()` can again be used to put the simulation results together given the nine  saved files (or three, if `multirow` is `TRUE` and `#SBATCH --array=1-3` were used instead). 

This type of hybrid approach is a middle ground between submitting the complete job (top of this vignette) and the `condition` + `replication` distributed load in the previous section, though has similar overhead + inefficiency issues as before (though less so, as the `array` jobs are evaluated independently). Moreover, if the row's take very different amounts of time to evaluate then this strategy can prove inefficient (e.g., the first two rows may take 2 hours to complete, while the third row may take 12 hours to complete; hence, the complete simulation results would not be available until the most demanding simulation conditions are returned!). Nevertheless, for moderate intensity simulations the above approach can be sufficient as each (batch) of simulation conditions can be evaluated independently across each `array` on the HPC cluster.

For more intense simulations, particularly those prone to time outs or other exhausted resources, the `runArraySimulation()` approach remains the recommended approach as the `max_RAM` and `max_time` fail-safes are more naturally accommodated within the replications, the jobs can be explicitly distributed given the anticipated intensity of each simulation condition, and the quality and reproducibility of multiple job submissions is easier to manage (see the FAQ section below). 

# Extra information (FAQs)

## Helpful Slurm commands

In addition to using `sbatch` to submit jobs, the following contains other useful Slurm commands.

```
sbatch <jobfile.slurm>  # submit job file to Slurm scheduler
squeue -u <username>    # what jobs are currently queued/running for a specific user
sshare -U <username>    # check the share usage for a specific user
scancel <jobid>         # cancel a specific job
scancel -u <username>   # cancel all queued and running jobs for a specific user
```

## My HPC cluster excution time/RAM is limited and terminates before the simulation is complete

This issue is important whenever the HPC cluster has mandatory time/RAM limits for the job submissions, where the array job may not complete within the assigned resources --- hence, if not properly managed, will discard any valid replication information when abruptly terminated. Unfortunately, this is a very likely occurrence, and is largely a function of being unsure about how long each simulation condition/replication will take to complete when distributed across the arrays (some conditions/replications will take longer than others, and it is difficult to be perfectly knowledgeable about this information beforehand) or how large the final objects will grow as the simulation progresses.

To avoid this time/resource waste it is **strongly recommended** to add a `max_time` and/or `max_RAM` argument to the `control` list (see `help(runArraySimulation)` for supported specifications), which are less than the Slurm specifications. These control flags will halt the `runArraySimulation()`/`runSimulation()` executions early and return only the complete simulation results up to this point. However, this will only work if these arguments are *non-trivially less than the allocated Slurm resources*; otherwise, you'll run the risk that the job terminates before the `SimDesign` functions have the chance to store the successfully completed replications. Setting these to around 90-95% of the respective `#SBATCH --time=` and `#SBATCH --mem=` inputs should, however, be sufficient in most cases.

```{r eval=FALSE}
# Return successful results up to the 11 hour mark, and terminate early 
#   if more than 3.5 GB of RAM are required to store the internal results
runArraySimulation(design=Design300, replications=replications,
                   generate=Generate, analyse=Analyse,
                   summarise=Summarise, iseed=iseed, arrayID=arrayID, 
                   dirname='mysimfiles', filename='mysim',
                   control=list(max_time="11:00:00", max_RAM="3.5GB"))   

```

Of course, if the session does time out early then this implies that the target `replications` will be missed on the first job submission batch. Therefore, and as is covered in the next section, a new job must be submitted to the scheduler that is mindful of the initial simulation history (particularly, the `.Random.seed` states). 

## Uploading array jobs related to previous array submissions

Related to early termination issue above is what to do about the missing replication information in the event that the complete set of replication information has not been collected. Obtaining the missing information clearly requires a second (or third) submission of the simulation job, though obviously only for the conditions where the collected replication results were problematic. Moreover, this has to be performed with care to avoid redundant random data generation strings, ultimately resulting in sub-optimal results.

To start, locate the simulation conditions in the aggregated result that do not meet the target replication criteria. This could be obtained via inspection of the aggregated results

```{r eval=FALSE}
Final <- SimCollect(files=dir())
Final
```

```
# A tibble: 3 × 8
      N    mean  median REPLICATIONS   SIM_TIME  COMPLETED               
<dbl>   <dbl>   <dbl>           <dbl>  <chr>     <chr>                   
1    10  9.9973  9.9934         9000   23.42s    Thu Apr  4 11:50:11 2024
2    20 10.007  10.015         10000   24.24s    Thu Apr  4 11:50:35 2024
3    30 10.003  10.007          8000   24.39s    Thu Apr  4 11:51:00 2024
```

or via the more informative (and less memory intensive) `SimCollect(..., check.only=TRUE)` flag.

```{r eval=FALSE}
Missed <- aggregate_simulations(files=dir(), check.only=TRUE)
Missed
```

```
# A tibble: 4 × 3
      N MISSED_REPLICATIONS TARGET_REPLICATIONS
  <dbl>               <int>               <int>
1    10                1000               10000
2    30                2000               10000
```

Next, build a new simulation structure containing only the missing information components.

```{r include=FALSE}
subDesign <- createDesign(N=c(10,30))
replications_missed <- c(1000, 2000)
```

```{r eval=FALSE}
subDesign <- subset(Missed, select=N)
replications_missed <- subset(Missed, select=MISSED_REPLICATIONS)
```

```{r}
subDesign
replications_missed
```

At this point, you can return to the above logic of organizing the simulation script job, distributing the information across as many array jobs as necessary to fill in the missing information. However, as before you must be *very careful* about the random number generators per row in `subDesign` **and** the original submission job. The fix in this case is straightforward as well: simply create a continuation from the previous logic, where the new elements are treated as additional rows in the resulting object as though they were part of the initial job submission. 

So, we now just glue on the new `subDesign` information to the original expanded version, though telling the scheduler to only evaluate these new rows in the `#SBATCH --array` specification (this is technically unnecessary, but is conceptually clear and keeps all simulation files and array IDs consistent). 

```{r}
rc <- 50
Design_left <- expandDesign(subDesign, rc) # smaller number of reps per array
Design_left

replications_left <- rep(replications_missed/rc, each=rc)
table(replications_left)

# new total design and replication objects
Design_total <- rbind(Design300, Design_left)
nrow(Design_total)
replications_total <- c(replications, replications_left)
table(replications_total)

# this *must* be the same as the original submission!
iseed <- 1276149341
```
Again, this approach simply expands the original simulation with 300 array jobs to
one with 400 array jobs as though the added structure was an intended part of the 
initial design (which is obviously wasn't, but is organized as such). 

Finally, in the `.slurm` submission file you no longer want to evaluate the first 1-300 cases,
as these `.rds` files have already been evaluated, and instead want to change the `--array` line from

```
#SBATCH --array=1-300
```
to
```
#SBATCH --array=301-400
```
Submit this job to compute all the missing replication information, which stores these files into the same working directory but with the new information stored as `mysim-301.rds` through `mysim-400.rds`. In this example, there will now be a total of 400 files that have been saved. Once complete, run 
```{r eval=FALSE}
# See if any missing still
SimCollect(files=dir(), check.only=TRUE)

# Obtain complete simulation results
Final <- SimCollect(files=dir())
```
one last time, which now reads in the complete set of 400 stored files instead of the previous 300, thereby obtaining the complete set of high-quality simulation results. Rinse and repeat if the same issue appears yet again on the second submission.
