---
title: "Parallel computing information"
author: "Phil Chalmers"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: false
    number_sections: true 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
vignette: >
  %\VignetteIndexEntry{Parallel computing information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r nomessages, echo = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.height = 5,
  fig.width = 5
)
options(digits=4)
par(mar=c(3,3,1,1)+.1)
```

# Cluster computing

`SimDesign` code may be released to a computing system which supports parallel cluster computations using
the industry standard Message Passing Interface (MPI) form. This simply
requires that the computers be setup using the usual MPI requirements (typically, running some flavor
of Linux, have password-less open-SSH access, IP addresses have been added to the `/etc/hosts` file
or `~/.ssh/config`, etc). More generally though, these resources are widely available through professional
organizations dedicated to super-computing.

To setup the R code for an MPI cluster the argument `MPI = TRUE` needs to be added to the 
`extra_options` list input, which wraps the appropriate MPI directives around `runSimulation`.
At this point the source files can be submitted using suitable BASH commands to execute the `mpirun` tool. For example,
```{r eval=FALSE}
library(doMPI)
cl <- startMPIcluster()
registerDoMPI(cl)
runSimulation(design=Design, replications=1000, filename='mysimulation',
    generate=Generate, analyse=Analyse, summarise=Summarise, extra_options = list(MPI=TRUE))
closeCluster(cl)
mpi.quit()
```
The necessary `SimDesign` files must be uploaded to the dedicated master node
so that a BASH call to `mpirun` can be used to distribute the work across slaves.
For instance, if the following BASH command is run on the master node then 16 processes
will be summoned (1 master, 15 slaves) across the computers named `localhost`, `slave1`,
and `slave2` in the ssh `config` file.

`mpirun -np 16 -H localhost,slave1,slave2 R --slave -f simulation.R`

A similar setup can also be used via the recently supported `future` interface (see below).

# Network computing

If you access have to a set of computers which can be linked via secure-shell (ssh) on the same LAN network then
Network computing (a.k.a., a Beowulf cluster) may be a viable and useful option.
This approach is similar to MPI computing approach
except that it offers more localized control and requires more hands-on administrative access to the master
and slave nodes. The setup generally requires that the master node
has `SimDesign` installed and the slave/master nodes have all the required R packages pre-installed
(Unix utilities such as `dsh` are very useful for this purpose). Finally,
the master node must have ssh access to the slave nodes, each slave node must have ssh access
with the master node, and a cluster object (`cl`) from the `parallel` package must be defined on the
master node.

Setup for network computing is generally more straightforward and controlled
than the setup for MPI jobs in that it only requires the specification of a) the respective
IP addresses within a defined R script, and b) the user name
(if different from the master node's user name. Otherwise, only a) is required).
However, on Linux I have found it is also important to include relevant information about the host names
and IP addresses in the `/etc/hosts` file on the master and slave nodes, and to ensure that
the selected port (passed to `makeCluster()`) on the master node is not hindered by a firewall.
As an example, using the following code the master node (primary) will spawn 7 slaves and 1 master,
while a separate computer on the network with the associated IP address will spawn an additional 6 slaves.
Information will be collected on the master node, which is also where the files
and objects will be saved using the `save` inputs.
```{r eval=FALSE}
library(parallel)
primary <- '192.168.2.1'
IPs <- list(list(host=primary, user='myname', ncore=8), list(host='192.168.2.2', user='myname', ncore=6))
spec <- lapply(IPs, function(IP) rep(list(list(host=IP$host, user=IP$user)), IP$ncore))
spec <- unlist(spec, recursive=FALSE)
cl <- makeCluster(master=primary, spec=spec, type = 'PSOCK')
Final <- runSimulation(..., cl=cl)
stopCluster(cl)
```
The object `cl` is passed to `runSimulation` on the master node
and the computations are distributed across the respective
IP addresses. Finally, it's usually good practice to use `stopCluster(cl)`
when all the simulations are said and done to release the communication between the computers,
which is what the above code shows.

Alternatively, if you have provided suitable names for each respective slave node, as well as the master,
then you can define the `cl` object using these instead (rather than supplying the IP addresses in
your R script). This requires that the master node has itself and all the slave nodes defined in the
`/etc/hosts` and `~/.ssh/config` files, while the slave nodes require themselves and the
master node in the same files (only 2 IP addresses required on each slave).
Following this setup, and assuming the user name is the same across all nodes,
the `cl` object could instead be defined with
```{r eval=FALSE}
library(parallel)
primary <- 'master'
IPs <- list(list(host=primary, ncore=8), list(host='slave', ncore=6))
spec <- lapply(IPs, function(IP) rep(list(list(host=IP$host)), IP$ncore))
spec <- unlist(spec, recursive=FALSE)
cl <- makeCluster(master=primary, spec=spec, type = 'PSOCK')
Final <- runSimulation(..., cl=cl)
stopCluster(cl)
```

A similar setup can also be used via the recently supported `future` interface (see below).

# Poor man's cluster computing for independent nodes

In the event that you do not have access to a Beowulf-type cluster (described in the section on
"Network Computing") but have multiple personal
computers then the simulation code can be manually distributed across each independent computer instead.

This simply requires passing a smaller value to the `replications` argument on each computer and later
aggregating the results using the `aggregate_simulations()` function.
For instance, if you have two computers available on different networks and wanted a total of 500 replications you
could pass `replications = 300` to one computer and`replications = 200` to the other along
with a `filename` argument (or simply saving the final objects as `.rds` files manually after
`runSimulation()` has finished). This will create two distinct `.rds` files which can be
combined later with the `aggregate_simulations()` function. The benefit of this approach over
MPI or setting up a Beowulf cluster is that computers need not be linked on the same network,
and, should the need arise, the temporary
simulation results can be migrated to another computer in case of a complete hardware failure by moving the
saved temp files to another node, modifying
the suitable `compname` input to `save_details` (or, if the `filename` and `tmpfilename`
were modified, matching those files accordingly), and resuming the simulation as normal.

Note that this is also a useful tactic if the MPI or Network computing options require you to
submit smaller jobs due to time and resource constraint-related reasons,
where fewer replications/nodes should be requested. After all the jobs are completed and saved to their
respective files, `aggregate_simulations()`
can then collapse the files as if the simulations were run all at once. Hence, `SimDesign` makes submitting
smaller jobs to super-computing resources considerably less error prone than managing a number of smaller
jobs manually.

# Using the `future` framework

The `future` framework (see `help(future, package = 'future')`) can also be used for distributing the
asynchronous function evaluations by changing the logical input in `runSimulation(..., parallel = TRUE/FALSE)` to the character vector `runSimulation(..., parallel = 'future')`, while the computation plan is pre-specified via `future::plan()`. For example, to initialize a local two-worker parallel processing computational plan one can use the follow:

```{r eval=FALSE}
library(future)
plan(multisession, workers = 2)

res <- runSimulation(design=Design, replications=1000, generate=Generate, 
                     analyse=Analyse, summarise=Summarise,
                     parallel = 'future')
```

The benefit of using the `future` framework is the automatic support of many distinct back-ends, such as, for instance, HPC clusters that control the distribution of jobs via Slurm or TORQUE (e.g., see the `future.batchtools` package). 

For progress reporting the `progressr` package is required and is intended as a wrapper around `runSimulation()`. Specifically, wrap the function `with_progress()` around `runSimulation()` after having specified the type of `handler()` to use, such as via the following. 

```{r eval=FALSE}
library(progressr)

# Rstudio style handler (if using RStudio)
handlers("rstudio")

# or using the cli package for terminal-based progress 
handlers('cli')

# See help(progressr) for additional options and details

# to use progressr, wrap/pipe inside with_progress() 
res <- with_progress(runSimulation(design=Design, replications=1000, generate=Generate, 
                     analyse=Analyse, summarise=Summarise,
                     parallel = 'future'))
```

Finally, when the parallel computations are complete be sure to manually reset the computation plan to free any workers via 

```{r eval=FALSE}
plan(sequential) # release workers
```


