%\VignetteIndexEntry{simTool}
%\VignetteEngine{knitr::knitr}
---
title: "SimDesign"
author: "Phil Chalmers"
date: "September 22, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r include=FALSE}
options(digits = 3)
```

# Simulation: Comparing two estimators and how effectively they recover population parameters

Item factor analysis (IFA) is a common method for determine the effectiveness and structure of items in
psychological tests. Two approaches currently exist: full-information IFA via method from the item response
theory framework, and limited-information IFA from the structural equation modeling framework. The question is,
under which circumstances should either method be used? Here we explore a simple one factor test structure
to determine which approach can recover the item parameters when varying test length and sample size.

### Define the conditions

Start by defining the conditions to be studied. Because we are interested in a number of slopes and intercepts
it is often convenient to save long strings of numbers to external files and read them in during the simulation.
Naturally, these could be included in the simulation source code directly, but often will take up a large amount
of space (could be hundreds of lines). Because these are fixed and known a prior, writing these to an external
file is a good strategy.

```{r}
sample_sizes <- c(100, 250, 500, 1000)
test_length <- c(10, 30)
Design <- expand.grid(sample_size=sample_sizes, test_length=test_length)

# save extra verbose information in an external file, and read it in later
set.seed(1234)
aux_info <- vector('list', 2)
names(aux_info) <- test_length
for(i in 1:2)
    aux_info[[i]] <- list(a = round(rlnorm(test_length[i], .2, .3), 2), 
                          d = round(rnorm(test_length[i], 0, .5), 2))
aux_info
saveRDS(aux_info, 'parameters.rds')
```

### Define the functions

As usual, define the functions of interest. Here we make sure that `lavaan` and `mirt` are loaded in when they
are required by calling a suitable `library()` call to make their functions available (required when 
running simulations in parallel).

```{r}
library(SimDesign)
# SimDesign_functions()
```

```{r}
Generate <- function(condition) {
    
    library(mirt) #for simdata() function
    nitems <- as.character(condition$test_length)
    N <- condition$sample_size
        
    #source in for convience (otherwise, these could be manually define in the source code)
    parameters <- readRDS('parameters.rds')
    a <- matrix(parameters[[nitems]]$a)
    d <- matrix(parameters[[nitems]]$d)
    
    dat <- as.data.frame(simdata(a=a, d=d, N=N, itemtype = 'dich'))
    return(dat)
}

Analyse <- function(dat, parameters, condition) {
    
    library(mirt)
    library(lavaan)
    nitems <- condition$test_length
    
    mod <- try(mirt(dat, 1L, verbose=FALSE), silent=TRUE)
    if(mod@converge != 1) stop('did not converge')
    cfs <- coef(mod, simplify = TRUE, digits = Inf)
    FIML_as <- cfs$items[,1L]
    FIML_ds <- cfs$items[,2L]
    
    lavmod <- paste0('F =~ ', paste0('NA*', colnames(dat)[1L], ' + '), 
                     paste0(colnames(dat)[-1L], collapse = ' + '),
                     '\nF ~~ 1*F')
    lmod <- try(sem(lavmod, dat, ordered = colnames(dat)))
    cfs2 <- coef(lmod) * 1.702 # scaling adjustment
    DWLS_as <- cfs2[1L:nitems]
    DWLS_ds <- -1 * cfs2[(1L:nitems) + nitems]

    return(c(FIML_as=unname(FIML_as), FIML_ds=unname(FIML_ds), 
             DWLS_as=unname(DWLS_as), DWLS_ds=unname(DWLS_ds)))
}

Summarise <- function(results, parameters_list, condition) {
    
    parameters <- readRDS('parameters.rds')
    nitems <- as.character(condition$test_length)
    pop_as <- matrix(parameters[[nitems]]$a)
    pop_ds <- matrix(parameters[[nitems]]$d)
    pop <- c(pop_as, pop_ds, pop_as, pop_ds)
    
    index <- 1:ncol(results)
    
    obt_bias <- sapply(index, function(ind, obs, pop) bias(obs[,ind], pop[ind]),
                       obs = results, pop = pop)
    obt_RMSE <- sapply(index, function(ind, obs, pop) RMSE(obs[,ind], pop[ind]),
                       obs = results, pop = pop)
    names(obt_bias) <- names(obt_RMSE) <- colnames(results)
    
    ret <- c(bias=obt_bias, RMSE=obt_RMSE)
    return(ret)
}
```

### Run the simulation

Because this simulation takes considerably longer it is recommended to pass the `save = TRUE` to temporarily
save results in case of power outages. Results can be continued by running the identical simulation code as the
initial run, and the function will automatically detect whether any temp files are available and resume the 
simulation at the previously saved location.

```{r cache=TRUE}
res10 <- runSimulation(subset(Design, test_length == 10), replications = 1000, verbose = FALSE, save = TRUE,
                       generate=Generate, analyse=Analyse, summarise=Summarise, parallel = TRUE)
```

```{r cache=TRUE}
res30 <- runSimulation(subset(Design, test_length == 30), replications = 1000, verbose = FALSE, save = TRUE,
                       generate=Generate, analyse=Analyse, summarise=Summarise, parallel = TRUE)
```

### Analyze the results

Sometimes, reshaping and indexing your output can be very helpful. Here we break the analysis into two parts,
though other strategies are certainly possible. Because analyzing simulations is a lot like analyzing empirical
data no one strategy may be the best; you have to use judgment.

#### Ten items

```{r}
# bias in slopes
names10 <- colnames(res10)
bias_as_fiml <- t(res10[,grepl('bias\\.', names10) & grepl('\\_as', names10) & 
                       grepl('FIML', names10)])
colnames(bias_as_fiml) <- sample_sizes
rownames(bias_as_fiml) <- aux_info[["10"]]$a

bias_as_dwls <- t(res10[,grepl('bias\\.', names10) & grepl('\\_as', names10) & 
                       grepl('DWLS', names10)])
colnames(bias_as_dwls) <- sample_sizes
rownames(bias_as_dwls) <- aux_info[["10"]]$a

(out <- list(FIML=bias_as_fiml, DWLS=bias_as_dwls))
sapply(out, colMeans)

# bias in intercepts
bias_ds_fiml <- t(res10[,grepl('bias\\.', names10) & grepl('\\_ds', names10) & 
                       grepl('FIML', names10)])
colnames(bias_ds_fiml) <- sample_sizes
rownames(bias_ds_fiml) <- aux_info[["10"]]$d

bias_ds_dwls <- t(res10[,grepl('bias\\.', names10) & grepl('\\_ds', names10) & 
                       grepl('DWLS', names10)])
colnames(bias_ds_dwls) <- sample_sizes
rownames(bias_ds_dwls) <- aux_info[["10"]]$d

(out <- list(FIML=bias_ds_fiml, DWLS=bias_ds_dwls))
sapply(out, colMeans)
```

DWLS does not appear to be less biased overall, and also does not appear to be come less biased as the 
sample size increases. Furthermore, the slopes seem to be systematically over estimated by a large amount 
in the DWLS as indicated by the large negative bias. FIML however does get better as the $N$ increases, 
and does not seem to be systematically biased in either the intercepts or slopes.

```{r}
# RMSE in slopes
RMSE_as_fiml <- t(res10[,grepl('RMSE\\.', names10) & grepl('\\_as', names10) & 
                       grepl('FIML', names10)])
colnames(RMSE_as_fiml) <- sample_sizes
rownames(RMSE_as_fiml) <- aux_info[["10"]]$a

RMSE_as_dwls <- t(res10[,grepl('RMSE\\.', names10) & grepl('\\_as', names10) & 
                       grepl('DWLS', names10)])
colnames(RMSE_as_dwls) <- sample_sizes
rownames(RMSE_as_dwls) <- aux_info[["10"]]$a

(out <- list(FIML=RMSE_as_fiml, DWLS=RMSE_as_dwls))
sapply(out, colMeans)

# RMSE in intercepts
RMSE_ds_fiml <- t(res10[,grepl('RMSE\\.', names10) & grepl('\\_ds', names10) & 
                       grepl('FIML', names10)])
colnames(RMSE_ds_fiml) <- sample_sizes
rownames(RMSE_ds_fiml) <- aux_info[["10"]]$d

RMSE_ds_dwls <- t(res10[,grepl('RMSE\\.', names10) & grepl('\\_ds', names10) & 
                       grepl('DWLS', names10)])
colnames(RMSE_ds_dwls) <- sample_sizes
rownames(RMSE_ds_dwls) <- aux_info[["10"]]$d

(out <- list(FIML=RMSE_ds_fiml, DWLS=RMSE_ds_dwls))
sapply(out, colMeans)
```

Concerning RMSE, FIML appeared to do less well in smaller sample sizes but become more effective than DWLS 
when $N >= 500$. This effect appeared to be more prevalent in the slopes rather than the intercepts. The 
conclusion about RMSE is that FIML tends to be more efficient in recovering parameters, but only after a certain
sample size is achieved.

Additionally, there appears to be an effect relating to the size of the parameters. Larger slopes tend to be recovered with more bias than slopes which are closer to 0. Had we obtained information about the parameter 
standard errors this would have been obvious as well, because the sampling variability for more extreme parameters tends to be much larger than more centralized parameters. 

#### Thirty items

```{r}
# bias in slopes
names30 <- colnames(res30)
bias_as_fiml <- t(res30[,grepl('bias\\.', names30) & grepl('\\_as', names30) & 
                       grepl('FIML', names30)])
colnames(bias_as_fiml) <- sample_sizes
rownames(bias_as_fiml) <- aux_info[["30"]]$a

bias_as_dwls <- t(res30[,grepl('bias\\.', names30) & grepl('\\_as', names30) & 
                       grepl('DWLS', names30)])
colnames(bias_as_dwls) <- sample_sizes
rownames(bias_as_dwls) <- aux_info[["30"]]$a

(out <- list(FIML=bias_as_fiml, DWLS=bias_as_dwls))
sapply(out, colMeans)

# bias in intercepts
bias_ds_fiml <- t(res30[,grepl('bias\\.', names30) & grepl('\\_ds', names30) & 
                       grepl('FIML', names30)])
colnames(bias_ds_fiml) <- sample_sizes
rownames(bias_ds_fiml) <- aux_info[["30"]]$d

bias_ds_dwls <- t(res30[,grepl('bias\\.', names30) & grepl('\\_ds', names30) & 
                       grepl('DWLS', names30)])
colnames(bias_ds_dwls) <- sample_sizes
rownames(bias_ds_dwls) <- aux_info[["30"]]$d

(out <- list(FIML=bias_ds_fiml, DWLS=bias_ds_dwls))
sapply(out, colMeans)
```

The same trend appeared in the 30 item simulation case, where again FIML was less biased in both intercepts and 
slopes than DWLS. However, it appears that in larger tests DWLS becomes even more biased when recovering slopes,
but less bias in recovering intercepts. 

```{r}
# RMSE in slopes
RMSE_as_fiml <- t(res30[,grepl('RMSE\\.', names30) & grepl('\\_as', names30) & 
                       grepl('FIML', names30)])
colnames(RMSE_as_fiml) <- sample_sizes
rownames(RMSE_as_fiml) <- aux_info[["30"]]$a

RMSE_as_dwls <- t(res30[,grepl('RMSE\\.', names30) & grepl('\\_as', names30) & 
                       grepl('DWLS', names30)])
colnames(RMSE_as_dwls) <- sample_sizes
rownames(RMSE_as_dwls) <- aux_info[["30"]]$a

(out <- list(FIML=RMSE_as_fiml, DWLS=RMSE_as_dwls))
sapply(out, colMeans)

# RMSE in intercepts
RMSE_ds_fiml <- t(res30[,grepl('RMSE\\.', names30) & grepl('\\_ds', names30) & 
                       grepl('FIML', names30)])
colnames(RMSE_ds_fiml) <- sample_sizes
rownames(RMSE_ds_fiml) <- aux_info[["30"]]$d

RMSE_ds_dwls <- t(res30[,grepl('RMSE\\.', names30) & grepl('\\_ds', names30) & 
                       grepl('DWLS', names30)])
colnames(RMSE_ds_dwls) <- sample_sizes
rownames(RMSE_ds_dwls) <- aux_info[["30"]]$d

(out <- list(FIML=RMSE_ds_fiml, DWLS=RMSE_ds_dwls))
sapply(out, colMeans)
```

Again the same trends appeared, however this time FIML become more efficient when $N >= 250$ in the slopes.



```{r include=FALSE}
system('rm *.rds')
```

