%\VignetteIndexEntry{SimDesign}
%\VignetteEngine{knitr::knitr}
---
title: "SimDesign"
author: "Phil Chalmers"
date: "September 22, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r include=FALSE}
options(digits = 3)
set.seed(1)
```

# Power-curve

Power-curves are useful when researchers are interested in the range of power effects when varying different population and sample properties. For instance, researchers may be interested in how power improves as the magnitude of the population mean difference for a t-test increases, or how quickly power improves when increasing the sample size ($N$).  

### Define the functions

```{r}
library(SimDesign)
#SimFunctions(comments = FALSE)

### Define design conditions. Here we want a finer range of values for plotting
Design <- expand.grid(mean_diff = seq(0, 1, by = .1), 
                      sample_size = c(10, 20, 30))

#--------------------------------------------------------------------

Generate <- function(condition, fixed_objects = NULL) {
    ret <- with(condition, rnorm(sample_size, mean_diff))
    ret
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(p = t.test(dat)$p.value)
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL) {
    ret <- EDR(results, alpha = .05)
    ret
}

#--------------------------------------------------------------------

### Run the simulation (using all available cores)
results <- runSimulation(Design, replications=10000, verbose=FALSE, parallel=TRUE,
                         generate=Generate, analyse=Analyse, summarise=Summarise, edit='none')
```

```{r}
# summary of simulation object
summary(results)

# print results
print(results)
```

A power curve is created by placing the detection rates on a $y$-axis and including different factors on the x-axis and other aesthetics (shading, colours, etc). Here we use `ggplot2` to construct suitable power-curves.

```{r fig.align='center'}
library(ggplot2)
ggplot(results, aes(mean_diff, p, colour=sample_size, group=sample_size)) + geom_point() + geom_line() +
    xlab('Mean Difference') + ylab('Detection Rate') + ggtitle('Empirical Power Curves') +
    scale_color_discrete('N')
```

Compare these results to analytical formulas, we can see that the two are essentially coinciding with good accuracy. 

```{r}
library(pwr)
pwr.t.test(d=0.5, n=30, sig.level=0.05, type="one.sample", alternative="two.sided")
subset(results, mean_diff == .5 & sample_size == 30)
pwr.t.test(d=0.5, n=10, sig.level=0.05, type="one.sample", alternative="two.sided")
subset(results, mean_diff == .5 & sample_size == 10)
```

# Power-curve function estimation

An alternative approach to estimating power curves is to obtain an approximate power function through parametric estimation. This involves estimating the same condition design while varying exactly one axillary condition, potentially with a smaller number of replications, and then fitting the observed proportions using a logistic response function. This is illustrated below. 

At best, however, this is an approximation and is not guaranteed to behave optimally (unlike the previous approach). Therefore, **for serious applications use the above approach instead**.

```{r}
library(SimDesign)
#SimFunctions(comments = FALSE)

### Find out parametric power curve for mean_diff = 0.3
Design <- expand.grid(mean_diff = 0.3, 
                      sample_size = 5:200)

#--------------------------------------------------------------------

Generate <- function(condition, fixed_objects = NULL) {
    ret <- with(condition, rnorm(sample_size, mean_diff))
    ret
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(reject = t.test(dat)$p.value < .05) 
    as.numeric(ret) #return 0/1 rather than FALSE/TRUE
}

#--------------------------------------------------------------------

### Run the simulation (note: no summarise function)
reps <- 50
results <- runSimulation(Design, replications=reps, verbose=FALSE, parallel=FALSE,
                         generate=Generate, analyse=Analyse, edit='none')
```

```{r}
res <- do.call(rbind, results)
head(res)
tail(res)

IV <- rep(Design$sample_size, each=reps)
(mod <- glm(res ~ IV, family = binomial))
plot(IV, plogis(predict(mod)), ylim = c(0,1), las=1,
	 type = 'l', xlab = 'Sample size', ylab = 'Power', main = 'Mean difference = 0.3')

# true power function
library(pwr)
out <- pwr.t.test(d=0.3, n=Design$sample_size, sig.level=0.05, type="one.sample", alternative="two.sided")
lines(Design$sample_size, out$power, col='red')
```

