%\VignetteIndexEntry{SimDesign}
%\VignetteEngine{knitr::knitr}
---
title: "SimDesign"
author: "Phil Chalmers"
date: "September 22, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r include=FALSE}
options(digits = 3)
set.seed(1)
```

# Power-curve

Power-curves are useful when researchers are interested in the range of power effects when varying different population and sample properties. For instance, researchers may be interested in how power improves as the magnitude of the population mean difference for a t-test increases, or how quickly power improves when increasing the sample size ($N$).  

### Define the functions

```{r}
library(SimDesign)
#SimFunctions(comments = FALSE)

### Define design conditions. Here we want a finer range of values for plotting
Design <- expand.grid(mean_diff = seq(0, 1, by = .1), 
                      sample_size = c(10, 20, 30))

#--------------------------------------------------------------------

Generate <- function(condition, fixed_objects = NULL) {
    ret <- with(condition, rnorm(sample_size, mean_diff)) #wrap in with() rather than using $
    ret
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    ret <- c(p = t.test(dat)$p.value)
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL) {
    ret <- EDR(results, alpha = .05)
    ret
}

#--------------------------------------------------------------------

### Run the simulation (using all available cores)
results <- runSimulation(Design, replications=10000, verbose=FALSE, parallel=TRUE,
                         generate=Generate, analyse=Analyse, summarise=Summarise, edit='none')
```

```{r}
# summary of simulation object
summary(results)

# print results
print(results)
```

A power curve is created by placing the detection rates on a $y$-axis and including different factors on the x-axis and other aesthetics (shading, colours, etc). Here we use `ggplot2` to construct suitable power-curves.

```{r fig.align='center'}
library(ggplot2)
ggplot(results, aes(mean_diff, p, colour=sample_size, group=sample_size)) + geom_point() + geom_line() +
    xlab('Mean Difference') + ylab('Detection Rate') + ggtitle('Empirical Power Curves') +
    scale_color_discrete('N')
```

Compare these results to analytical formulas, we can see that the two are essentially coinciding with good accuracy. 

```{r}
library(pwr)
pwr.t.test(d=0.5, n=30, sig.level=0.05, type="one.sample", alternative="two.sided")
subset(results, mean_diff == .5 & sample_size == 30)
pwr.t.test(d=0.5, n=10, sig.level=0.05, type="one.sample", alternative="two.sided")
subset(results, mean_diff == .5 & sample_size == 10)
```

