%\VignetteIndexEntry{SimDesign}
%\VignetteEngine{knitr::knitr}
---
title: "SimDesign"
author: "Phil Chalmers"
date: "September 22, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r include=FALSE}
options(digits = 3)
```

# Simulation with mirtCAT

Example simulation for computerized adaptive testing with the `mirtCAT` package. Use of `SimDesign` for this
package demonstrates a different type of parallelization, whereby the work is distributed within the `analysis`
step rather than across all replications. This showcases how within-function parallelization can be utilized for 
code which has already been optimized for simulations, but still needs an 'outer-shell' to collect the simulation
results. Currently requires `mirtCAT` version 1.1.4 or higher.

This study explored two major design conditions: size of the item bank, and the magnitude of the discrimination parameter. Just for didactic purposes the discrimination parameters were all set equal, however real studies will
call for more intimate control of the properties of the item bank.

```{r}
library(SimDesign)
library(mirtCAT)
library(parallel)

Design <- createDesign(nitems = c(100, 300, 500),
					   discrimination = c(.5, 1.5))
cl <- makeCluster(detectCores())
fo <- list(cl=cl, 
		   design=list(min_SEM = .25, min_items = 10, max_items = 50))

#-------------------------------------------------------------------

Generate <- function(condition, fixed_objects = NULL) {
	Attach(condition)
    Theta <- matrix(rnorm(1000))
	a <- matrix(rep(discrimination, nitems))
	d <- rnorm(nitems)
	pars <- data.frame(a1 = a, d = d, g = 0.2)
	mirt_object <- generate.mirt_object(pars, '3PL')
	responses <- generate_pattern(mirt_object, Theta = Theta)
    list(mirt_object=mirt_object, responses=responses)
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
	ret <- with(fixed_objects, 
				mirtCAT(mo = dat$mirt_object, local_pattern = dat$responses,
						start_item = 'MI', method = 'EAP',
						criteria = 'MI', design = design, cl = cl))
    ret
}

#-------------------------------------------------------------------

# only 1 replication
res <- runSimulation(design=Design, replications=1, seed = 1:nrow(Design),
                     generate=Generate, analyse=Analyse, fixed_objects=fo, debug='none')
```


```{r}
res

# person 100 from the first condition
plot(res$`nitems=100; discrimination=0.5`[[1]][[100]], SE = 1.96)

# person 50 from the last condition
plot(res$`nitems=500; discrimination=1.5`[[1]][[50]], SE = 1.96)

diff <- lapply(res, function(x){
	sapply(x[[1]], function(y){
		so <- summary(y)
		so$final_estimates[1,] - so$true_thetas
	})
})

nitems_answered <- lapply(res, function(x){
	sapply(x[[1]], function(y){
		so <- summary(y)
		length(so$items_answered)
	})
})

meta_stats <- data.frame(Design, bias=sapply(diff, bias), RMSE=sapply(diff, RMSE), 
						 nitems_answered=sapply(nitems_answered, mean), row.names = NULL)
meta_stats
```

From this simple summary we can see that not only did the conditions with larger discrimination parameters
provide better accuracy, but the CATs were actually terminated before the maximum 50 item cutoff. 
