%\VignetteIndexEntry{SimDesign}
%\VignetteEngine{knitr::knitr}
---
title: "Flora and Curran (2004) simulation"
author: "Phil Chalmers"
output:
  html_document:
    number_sections: yes
    toc: yes
---

```{r include=FALSE}
options(digits = 3)
cache_ <- TRUE
```

Simulation from:

- Flora, D. B. & Curran, P. J. (2004). An Empirical Evaluation of Alternative Methods of Estimation for Confirmatory Factor Analysis With Ordinal Data. Psychological Methods, 9, 466-491

# User-defined functions

User-defined helpter function for the generate-analyse steps to generate appropriate syntax for `lavaan` (could be placed in an external .R file and sourced in prior to running `runSimulation()` instead).

```{r}
#' @param J number of variables
#' @param analyse logical; is syntax being used to data generation or analysis?
#' @examples
#' cat(genLavaanSyntax(factors=1, indicators=10))
#' cat(genLavaanSyntax(factors=2, indicators=10))
#' cat(genLavaanSyntax(factors=1, indicators=10, analyse = TRUE))
#' cat(genLavaanSyntax(factors=2, indicators=10, analyse = TRUE))
#'
genLavaanSyntax <- function(factors, indicators, analyse = FALSE){
    ret <- if(factors == 1){
        if(analyse){
            paste0(paste0('f1 =~ NA*x1 + ', paste0(paste0('x', 2:indicators),
                                                   collapse=' + ')),
                   '\nf1 ~~ 1*f1')
        } else {
            paste0(paste0('f1 =~ ', paste0(rep(.7, indicators), 
                                           paste0('*x', 1:indicators),
                                           collapse=' + '), ' \n'),
                   paste0(sprintf('x%s ~~ 0.51*x%s', 1:indicators, 1:indicators), 
                          collapse=' \n'))
        }
    } else if(factors == 2){
        if(analyse){
            paste0(paste0('f1 =~ NA*x1 + ',
                          paste0(paste0('x', 2:indicators), collapse=' + ')),
                   paste0(sprintf('\nf2 =~ NA*x%s + ', indicators+1),
                          paste0(paste0('x', 2:indicators + indicators), collapse=' + ')),
                   '\nf1 ~~ 1*f1 \nf2 ~~ 1*f2 \nf1 ~~ f2')
        } else {
            paste0(paste0('f1 =~ ', paste0(rep(.7, indicators), 
                                           paste0('*x', 1:indicators),
                                           collapse=' + '), ' \n'),
                   paste0('f2 =~ ', paste0(rep(.7, indicators), 
                                           paste0('*x', 1:indicators + indicators),
                                           collapse=' + '), ' \n'),
                   'f1 ~~ .3*f2 \n',
                   paste0(sprintf('x%s ~~ 0.51*x%s', 1:indicators, 1:indicators), 
                          collapse=' \n'))
        }
    } else stop('factors input is incorrect')
    ret
}
```

# Simulation code

```{r eval=FALSE}
library(SimDesign)

Design <- createDesign(N = c(100, 200, 500, 1000),
                       categories = c(2, 5),
                       skewness_kurtosis = list(c(0, 0), c(.75, 1.75), c(.75, 3.75),
                                                c(1.25, 1.75), c(1.25, 3.75)),
                       factors = c(1, 2),
                       indicators = c(5, 10),
                       estimator = c('WLSMV', 'WLS'),
                       # remove poorly converging combinations 
                       subset = !(estimator == 'WLS' & N %in% c(100, 200) &
                                      factors == 2 & indicators == 10))

######################################################

Generate <- function(condition, fixed_objects = NULL) {
    Attach(condition)
    syntax <- genLavaanSyntax(factors=factors, indicators=indicators)
    cdat <- simulateData(syntax, model.type='cfa', sample.nobs=N,
                         skewness=skewness_kurtosis[1L],
                         kurtosis=skewness_kurtosis[2L])
    tau <- if(categories == 5)
        c(-1.645, -0.643, 0.643, 1.645) else 0
    # data generation fix described in Flora's (2002) unpublished dissertation
    if(categories == 5 && all(skewness_kurtosis == c(1.25, 1.75)))
        tau[1] <- -1.125
    dat <- apply(cdat, 2, function(x, tau){
        dat <- numeric(length(x))
        for(i in 1:length(tau))
            dat[x > tau[i]] <- i
        dat
    }, tau=tau)
    # throw error if number of categories not correct
    if(!all(apply(dat, 2, function(x) length(unique(x))) == categories))
        stop('Number of categories generated is incorrect')
    dat
}

Analyse <- function(condition, dat, fixed_objects = NULL) {
    Attach(condition)
    syntax <- genLavaanSyntax(factors=factors, indicators=indicators, analyse=TRUE)
    mod <- cfa(syntax, dat, ordered=colnames(dat), estimator=estimator)

    # check that model and coefficients are reasonable
    if(!lavInspect(mod, 'converged')) stop('Model did not converge')
    pick_lambdas <- matrix(TRUE, indicators*factors, factors)
    if(factors == 2)
        pick_lambdas[(indicators+1):(indicators*3)] <- FALSE
    cfs <- lavInspect(mod, what="std")$lambda[pick_lambdas]
    if(any(cfs > 1 | cfs < -1))
        stop('Model contains Heywood cases')
    if(factors > 2 && abs(lavInspect(mod, what="std")$psi[2,1]) >= 1)
        stop('Latent variable psi matrix not positive definite')

    # extract desired results
    fit <- fitMeasures(mod)
    ses <- lavInspect(mod, what="se")$lambda[pick_lambdas]
    fitstats <- if(estimator == 'WLS') fit[c('chisq', 'df', 'pvalue')]
    else if(estimator == 'WLSMV') fit[c('chisq.scaled', 'df.scaled', 'pvalue.scaled')]
    names(fitstats) <- c('chisq', 'df', 'pvalue')
    phi21 <- if(factors == 2)
        lavInspect(mod, what="std")$psi[1,2] else NULL

    ret <- c(fitstats, mean_ses=mean(ses), lambda=cfs, phi21=phi21)
    ret
}

Summarise <- function(condition, results, fixed_objects = NULL) {
    # model parameters
    lambdas <- results[ , grepl('lambda', colnames(results))]
    pool_mean_lambdas <- mean(apply(lambdas, 2, mean))     # Equation 10
    pool_SD_lambdas <- sqrt(mean(apply(lambdas, 2, var)))  # Equation 11
    RB_phi21 <- if(condition$factors == 2)
        bias(results$phi21, parameter=.3, type='relative', percent=TRUE) else NULL
    mean_se <- mean(results$mean_ses)

    # goodness-of-fit
    edr_05 <- EDR(results$pvalue, alpha = .05)
    mean_X2 <- mean(results$chisq)
    sd_X2 <- sd(results$chisq)
    RB_X2 <- bias(results$chisq, parameter=results$df, type='relative',
                  percent=TRUE, unname=TRUE)

    ret <- c(mean_X2=mean_X2, sd_X2=sd_X2, edr_05=edr_05,
             pool_mean_lambdas=pool_mean_lambdas,
             pool_SD_lambdas=pool_SD_lambdas, mean_se=mean_se,
             RB_X2=RB_X2, RB_phi21=RB_phi21)
    ret
}

######################################################

# run simulation
res <- runSimulation(design=Design, replications=500, generate=Generate,
                     analyse=Analyse, summarise=Summarise, max_errors=100,
                     packages='lavaan', parallel=TRUE, save=TRUE,
                     filename='FloraCurran2004', save_results=TRUE)
res
```

```{r echo=FALSE}
library(SimDesign)
res <- readRDS("../extras/FloraCurran2004/FloraCurran2004.rds")
res
```

# Extras

Recreate Table 5. 

```{r}
library(dplyr)

# separate s and k values in skewness_kurtosis, create % variable, 
#    select only 5 categories, rename columns to match
res %>% mutate(s = sapply(res$skewness_kurtosis, function(x) x[1]),
               k = sapply(res$skewness_kurtosis, function(x) x[2]),
               "% reject" = edr_05 * 100) %>% 
    filter(categories == 5) %>% 
    rename(M='mean_X2', SD='sd_X2', RB='RB_X2') -> res_5

# construct Table 5 information from Flora and Curran (2004)
res_5 %>%
    filter(indicators == 10, factors == 1, estimator == 'WLS') %>%  
    select(N, s, k, M, SD, RB, "% reject") -> WLS
res_5 %>% 
    filter(indicators == 10, factors == 1, estimator == 'WLSMV') %>%  
    select(N, s, k, RB, "% reject") -> WLSMV
full_join(WLS, WLSMV, by = c("N", "s", "k"), suffix=c('', '.DWLS')) %>% 
    arrange(N, s, k) -> tab5
pander::pander(tab5)
```

Recreate Figure 6. 

```{r}
library(ggplot2)

# generate summary data for Figure 6 in Flora and Curran (2004)
res_5 %>%
  filter(indicators == 5) %>%
  mutate(TypeI = edr_05 * 100,
         est_fact = factor(estimator):factor(factors)) %>%
  group_by(N, est_fact) %>%
  summarise(mean_TypeI = mean(TypeI)) -> fig6dat

labels <- c("Model 1 (full WLS estimation)", "Model 3 (full WLS estimation)",
            "Model 1 (robust WLS estimation)", "Model 3 (robust WLS estimation)")

# draw graphic using the ggplot2 package
ggplot(fig6dat,
       aes(x=factor(N), y=mean_TypeI, linetype=est_fact,
           shape=est_fact, group=est_fact)) +
  geom_line() + geom_point(size=4) +
  geom_hline(yintercept=5, colour='red', linetype='dashed') +
  xlab("Sample Size") + ylab("Type I Error Rate") +
  scale_linetype_discrete(labels=labels) +
  scale_shape_discrete(labels=labels) +
  ylim(0, 80) + theme_bw() +
  theme(legend.title=element_blank(),
        legend.position = c(.8, .7),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())
```





